Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-03-20 draftfinalreport_v2.Rnw
change to draft

Previous measles analyses, including two in New Zealand by Prof. Roberts, estimated the interruption of measles virus transmission can be achieved by herd immunity when approximately 95 percent of the population is homogeneously immune to measles %\citep{roberts0,roberts4}.
Thus, while New Zealand immunisation activities have led to measles outbreaks becoming less frequent, with decreasing numbers of cases, outbreaks still occur (as described above). Current estimates suggest that approximately 85 to 90 percent of the population is immune to measles (see \autoref{sub:immunestat}), thus the reasons for the ongoing outbreaks are likely due to overall population immunity being less than 95 percent and there being pockets of susceptible, non-immune population remaining. Since 2009, all the outbreaks in New Zealand have been linked to infections acquired (imported) from overseas, though previous work suggests these outbreaks still largely affect school-aged children (especially young adults) and children under two years of age. Those under two year olds are thought to be consistently among the most affected age groups because the first of two doses of measles, mumps and rubella vaccine (MMR) is not due until fifteen months.

@@

<<echo=FALSE,include=FALSE,results=hide>>=
## get data
# set wd
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)
test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 15,height=12>>=
par(mfrow=c(1,1))
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="Frequency",cex.lab=2)
hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=2)
@
\end{center}
\caption{Numbers and age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
## plot cases / age class
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = test , FUN=sum)
#caseyr
@

<<echo=FALSE,include=FALSE,results=hide>>=
require(reshape2)  # this is the library that lets one flatten out data
require(ggplot2)   # plotting library
popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
         rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
@

<<echo=FALSE,include=FALSE,results=hide>>=
## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
colnames(naive)<-c("AgeInYears","Naive","Cases")

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
library(plyr)
testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

testv$VC<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,0,
                 ifelse(testv$Dose1Mths >= 0 & is.na(testv$Dose2Mths) == T,1,2))
AgeVac<-table(testv$VC,testv$AgeInYears)
@

<<echo=FALSE,include=FALSE,results=hide>>=
row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)
AgeInYears<-(as.numeric(rownames(AgeVac)))
AgeVac<-cbind(AgeVac,AgeInYears)       
AgeInYears<-(as.numeric(rownames(pop)))
pop<-cbind(pop,AgeInYears)
colnames(pop)<-c("Population","AgeInYears")
AgeV<-merge(pop,AgeVac,by="AgeInYears",all=T)
# head(AgeV)
#AgeVac<-t(AgeVac)
@

The Prioritized Ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MLA); Other Ethnicity; Residual Categories. For the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity (see \emph{None}, \autoref{table:none}). 
<<echo=FALSE,include=FALSE,results=hide>>=
## regression analyses - measles
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = test , FUN=sum)
caseyr
test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
                                     "Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
                              ifelse (test$RptYear == "2008",test$NZDep06,
                              ifelse (test$RptYear == "2009",test$NZDep06,
                              ifelse (test$RptYear == "2010",test$NZDep06,
                              ifelse (test$RptYear == "2011",test$NZDep06,
                              ifelse (test$RptYear == "2012",test$NZDep06,
                              ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
head(testt)
tail(testt)
str(testt)
testt$NZDep<-as.factor(testt$NZDep-1)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
                 data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]
@

\begin{table}[hbtp]
\footnotesize
<<results=tex,echo=FALSE>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(ttyr, file="",table.env=FALSE,rowname=NULL)
@
\caption{Absolute number of measles cases with no ethnicity recorded for specific age and socio-economic deprivation (NZDep) categories from 2007-2014.}
\label{table:none}
\end{table}

<<echo=FALSE,include=FALSE,results=hide>>=
### denominator data
# setwd("~/Massey_2014/measles/data")
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = test , FUN=sum)
caseyr
test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
                                     "Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
                              ifelse (test$RptYear == "2008",test$NZDep06,
                              ifelse (test$RptYear == "2009",test$NZDep06,
                              ifelse (test$RptYear == "2010",test$NZDep06,
                              ifelse (test$RptYear == "2011",test$NZDep06,
                              ifelse (test$RptYear == "2012",test$NZDep06,
                              ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
head(testt)
tail(testt)
str(testt)
testt$NZDep<-as.factor(testt$NZDep-1)

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
popn
pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
                                 "Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
                                 "Dep9"="6-10","Dep10"="6-10"));

tp$cases <- 0
cases <- matrix(0, length(tp$merge),1)
rownames(cases) <- tp$merge
cases[ttyr$merge,] <- ttyr$Cases
tp$cases <- cases
PerCapita<-(tp$cases/tp$Popn)
PerCap<-cbind(tp[,c(1:4,6)],round(PerCapita*10000,4))
colnames(PerCap)<-c("NZDep","Age","Ethnicity","Population","Cases","Per capita")
@

<<echo=FALSE,include=FALSE,results=hide>>=

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)
test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 15,height=12>>=
par(mfrow=c(1,1))
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="Frequency",cex.lab=2)
hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=2)
@
\end{center}
\caption{test}
%\label{fig:test}
\end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
## plot cases / age class
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

require(reshape2)  # this is the library that lets one flatten out data
require(ggplot2)   # plotting library
popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))

@@

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
library(plyr)
testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

@@

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
# test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
#tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

@@

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
head(testt)
tail(testt)
str(testt)
testt$NZDep<-as.factor(testt$NZDep-1)

@@

#tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
#"5"="1-5","6"="6-10","7"="6-10","8"="6-10",
#"9"="6-10","10"="6-10"));

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]

# This chunk uses "results=tex"
library(Hmisc)
latex(ttyr, file="",table.env=FALSE,rowname=NULL)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)
library(reshape)
library(reshape2)
library(plyr)
library(pscl)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

>>>>

Previous measles analyses, including two in New Zealand by Prof. Roberts, estimated the interruption of measles virus transmission can be achieved by herd immunity when approximately 95 percent of the population is homogeneously immune to measles \citep{roberts0,roberts4}.
Thus, while New Zealand immunisation activities have led to measles outbreaks becoming less frequent, with decreasing numbers of cases, outbreaks still occur (as described above). Current estimates suggest that approximately 85 to 90 percent of the population is immune to measles (see \autoref{sub:immunestat}), thus the reasons for the ongoing outbreaks are likely due to overall population immunity being less than 95 percent and there being pockets of susceptible, non-immune population remaining. Since 2009, all the outbreaks in New Zealand have been linked to infections acquired (imported) from overseas, though previous work suggests these outbreaks still largely affect school-aged children (especially young adults) and children under two years of age. Those under two year olds are thought to be consistently among the most affected age groups because the first of two doses of measles, mumps and rubella vaccine (MMR) is not due until fifteen months.

@@

<<echo=FALSE,include=FALSE,results=hide>>=
library(adehabitat)
library(classInt)
library(ggplot2)   # plotting library
library(grid)
library(gridExtra)
library(ggmap)
library(Hmisc)
library(maps)       # Provides functions that let us plot the maps
library(mapdata)    # Contains the hi-resolution points that mark out the countries.
library(maptools)
library(mapproj)
library(plyr)
library(pscl)
library(RColorBrewer)
library(rworldmap)
library(reshape)
library(reshape2)
library(rgeos)
library(rgdal)
library(raster)
library(sp)
library(utils)
library(xtable)
#library(SDMTools)
#gpclibPermit()

# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
# names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)
time_yr<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = time_yr , FUN=sum)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
         rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
naive<-round(pop-(pop*impop))

@@

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = test , FUN=sum)
test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
                                     "Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth

test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
                              ifelse (test$RptYear == "2008",test$NZDep06,
                              ifelse (test$RptYear == "2009",test$NZDep06,
                              ifelse (test$RptYear == "2010",test$NZDep06,
                              ifelse (test$RptYear == "2011",test$NZDep06,
                              ifelse (test$RptYear == "2012",test$NZDep06,
                              ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
testt$NZDep<-as.factor(testt$NZDep-1)

testtable<-aggregate( cbind( DiseaseName) #+ as.numeric(SurvWeek))
                      ~ NZDep+AgeInYears+EthnicityPrioritised + RptYear, 
                 data = testt , FUN=sum)

colnames(testtable)<-c("NZDep","Age","Ethnicity","Year","Cases")

testtable$NZDep<-tnzd

tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
                 data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
                 data = testtable , FUN=sum)
denom<-read.csv("NZDep2006Denominators.csv",header=T)
denom<-denom[,-c(9)]

denom<-denom[,1:18]
row.has.na <- apply(denom, 1, function(x){any(is.na(x))})

denomt <- denom[!row.has.na,]

denomt<-denomt[denomt$Sex_code==99,]
denomt$Eth_Level<-as.factor(denomt$Eth_Level)
denomt<-denomt[denomt$Eth_Level %in% c("24","21","66",'22','77'),]
dm<-cbind(denomt[,c(3,6,9:18)])

dm<-dm[dm$Age_Label %in% c("A: 0- 5 yrs", "A: 6-17 yrs","A:18-24 yrs","A:25-64 yrs","A:65+ yrs"),]
dage<-revalue(dm$Age_Label, c("A: 0- 5 yrs"="<6", "A: 6-17 yrs"="6-17","A:18-24 yrs"="18-24","A:25-64 yrs"="25+","A:65+ yrs"="25+"));
dm$Age_Label<-dage
colnames(dm)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")

datadm<-dm[dm$Age=="<6",]
dataA<-(rbind(datadm,datadm)/2)
dataA<-round(dataA[,3:12])
dataA<-cbind(c(rep("0-2",5),rep("3-5",5)),rep(datadm$Ethnicity[1:5],2),dataA)
colnames(dataA)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")
ddm<-dm[!dm$Age=="<6",]
dm<-rbind(dataA,ddm)
dm$Ethnicity <- factor(dm$Ethnicity)
deth<-revalue(dm$Ethnicity, c("Asian (Prioritised)"="Asian","European (NZ European and Other European)"="European",
                              "Maori (Prioritised)"="Maori","MELAA"="MLA","Pacific People (Prioritised)"="Pacific"));
dm$Ethnicity<-deth
head(dm)

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
                                 "Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
                                 "Dep9"="6-10","Dep10"="6-10"));

popn$NZDep<-pnzd
tp<-aggregate( cbind(Popn) ~ NZDep+Age+Ethnicity, 
               data = popn , FUN=sum)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))

naive<-round(pop-(pop*impop))

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

@@

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

# test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
#tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth

test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))

testt$NZDep<-as.factor(testt$NZDep-1)

testtable$NZDep<-tnzd

tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]
### denominator data

data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)

data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

@@

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))

testt$NZDep<-as.factor(testt$NZDep-1)

@@
